The matter of the translation and rotation of the spherical harmonic functions is treated in its entirety by \textsc{Gumerov} \& \textsc{Duraiswami} in a plethora of articles on the topic of multipole expansions and their swift evaluation by the fast multipole method.
Chief among them is the paper wherein these rotations and translations are derived in detail,%
\footnote{%
  \cite{gumerov2003recursions} \textsc{Gumerov} \& \textsc{Duraiswami} (2003)
}
their book on the fast multipole method applied to the \textsc{Helmholtz} equation in particular serving as a readable alternative.%
\footnote{%
  \cite{gumerov2004fast} \textsc{Gumerov} \& \textsc{Duraiswami}, \S3
}
We shall however address first the matter of the result of the preceding text.
We shall insert into equation \eqref{eq:hypersphere_green} the relation \eqref{eq:gegenbauer-legendre} for the threedimensional case, yielding
%
\begin{equation}\label{eq:Green_translation}
  \green(\xvec; \bzhe) = \frac{1}{4\pi\absl{\bzhe}} \sum_{k = 0}^{\infty} \varepsilon^k \legendre_k(\cos{\alpha}).
\end{equation}

\noindent
We shall exploit the fact that the \textsc{Gegenbauer} functions evaluated at the inner product between two higherdimensional vectors can be written as a superposition of products of the spherical harmonics.%
\footnote{%
  \cite{avery2018hyperspherical} \textsc{Avery} \& \textsc{Avery}, \S3.2
}
This result simplifies further to what the literature on fast multipole calculations often labels the addition theorem for three dimensions, and is given as follows.

\begin{theorem}
  Let $\legendre_n$ denote the \textsc{Legendre} polynomial of order $n$, $\spherical_{n}^{m}$ be a spherical harmonic, and $\xvec_1$, $\xvec_2$ be points on the unit two-sphere, such that $\xvec_i = (\cos{\alpha_{i}^{1}}, \sin{\alpha_{i}^{0}} \sin{\alpha_{i}^{1}}, \cos{\alpha_{i}^{0}} \sin{\alpha_{i}^{1}})$.
  Then,
  \begin{equation}
    \legendre_n(\cos{\alpha}) = \frac{4\pi}{2n+1} \sum_{m = -n}^{n} \spherical_{n}^{m}(\alpha_{1}^{0}, \alpha_{1}^{1}) {\spherical_n^m}^{\dag}(\alpha_{2}^{0}, \alpha_{2}^{1}),
  \end{equation}
  where $\alpha$ is the angle between the points.
\end{theorem}

\noindent
The coëfficient is due to the normalization condition imposed upon the spherical harmonics.%
\footnote{%
  \cite{epton1995multipole} \textsc{Epton} \& \textsc{Dembart} (1995), eq.2.19
}
Note also the conjugate property that ${\Upsilon^m}^{\dag} = \Upsilon^{-m}$.
We may now address the translation directly.
It shall promptly become apparent that we seek local and distal expansions of the \textsc{Green} function.
Relative to some new origin $O^{\prime}$ with vector $\ovec$ from the absolute origin, we write the point of evaluation $\bzhe^{\prime} = \bzhe - \ovec$, the same with some source point $\xvec^{\prime}$.
The local and distal expansions shall respectively regular and singular at this new origin.
In particular, we shall have that%
\footnote{%
  \cite{gumerov2013efficient} \textsc{Gumerov} \& \textsc{Duraiswami} (2013), eq.25,28
}
%
\begin{equation}
  \green(\xvec; \bzhe) = \frac{1}{4\pi} \sum_{n = 0}^{\infty} \sum_{m = -n}^{n} \regular_{n}^{m}(-\xvec^{\prime}) \singular_{n}^{m}(\bzhe^{\prime}), \qquad
  \varepsilon^{\prime} < 1;
\end{equation}
%
\begin{subequations}
  \begin{equation}\tag{\theequation{a, b}}
    \regular_n^m (\bzhe^{\prime}) = \frac{{(-1)}^n i^{\absl{m}} {\absl{\bzhe^{\prime}}}^n}{c(n + \absl{m})!} \Upsilon_n^m(\bzhe^{\prime}), \qquad
    \singular_n^m(\bzhe^{\prime}) = \frac{(n - \absl{m})!}{i^{\absl{m}} c \absl{\bzhe^{\prime}}^{n+1}} \Upsilon_n^m(\bzhe^{\prime}).
  \end{equation}
\end{subequations}

\noindent
The arguments of the spherical harmonics is understood to be the angular part.
Note the interpretation now of switching the arguments of the \textsc{Green} function---we are explicitly changing which of the two variables contribute regularly, and \emph{vice versa} for thew singularity.
The act of translating either harmonic by the regular and singular functions is to be interpreted as the reëxpansion of that harmonic locally or distally.
In more concrete terms, we consider the position $\ovec_i$ of some origin one wishes to translate from, and the position $\ovec_j$ of some new origin one wishes to translate to.
The displacement thereof we denote $\dvec = \ovec_j - \ovec_i$, such that the translations are given as follows.%
\footnote{%
  \cite{gumerov2004fast} \textsc{Gumerov} \& \textsc{Duraiswami}, ch.5
}
%
\begin{equation}\label{eq:R2R}
  \regular_n^m (\dvec + \bzhe_j^{\prime}) = \sum_{n^{\prime} = 0}^{\infty} \sum_{m^{\prime} = -n^{\prime}}^{n^{\prime}} \regular_{n - n^{\prime}}^{m - m^{\prime}} (\dvec) \regular_{n^{\prime}}^{m^{\prime}} (\bzhe_{j}^{\prime})
\end{equation}
%
\begin{equation}\label{eq:S2R}
  \singular_n^m (\dvec + \bzhe_j^{\prime}) = \sum_{n^{\prime} = 0}^{\infty} \sum_{m^{\prime} = -n^{\prime}}^{n^{\prime}} \singular_{n + n^{\prime}}^{m - m^{\prime}} (\dvec) \regular_{n^{\prime}}^{m^{\prime}} (\bzhe_{j}^{\prime})
\end{equation}
%
\begin{equation}\label{eq:S2S}
  \singular_n^m (\dvec + \bzhe_j^{\prime}) = \sum_{n^{\prime} = 0}^{\infty} \sum_{m^{\prime} = -n^{\prime}}^{n^{\prime}} \regular_{n^{\prime} - n}^{m - m^{\prime}} (\dvec) \singular_{n^{\prime}}^{m^{\prime}} (\bzhe_{j}^{\prime})
\end{equation}

\noindent
For the translation of a singular function to one of regularity as with equation \eqref{eq:S2R}, we must evidently have that $\absl{\bzhe_j^{\prime}} < \absl{\dvec}$, and conversely for the translation of a singular expansion to another singular expansion as in equation \eqref{eq:S2S}, requiring that $\absl{\bzhe_j^{\prime}} > \absl{\dvec}$.
Recall the magnitudinal ordering of the indices denoting the azimuthal \textsc{Fourier} mode $m$, and the angular degree $n$.
The multiplication of this azimuthal term in the spherical harmonic must yield the proper azimuthal index, whence the subtraction of the indices in the intermediary expansion about $\dvec$.
The index in angular degree is similar in that regard, as the total degree of polynomials must additively coïncide due to the additive property of exponents.
We shall later discuss these translations in terms of the translations needed for the fast multipole method, in the context thereof the customarily named fast multiple kernels of equations \eqref{eq:R2R}--\eqref{eq:S2S} are bestowed the following notation.
%
\begin{subequations}\label{eq:translation_notation}
  \begin{equation}\tag{\theequation{a, b, c}}
    {\big[ \regular\vert\regular \big]}_{n^{\prime}n}^{m^{\prime}m} = \regular_{n - n^{\prime}}^{m - m^{\prime}}, \qquad
    {\big[ \singular\vert\regular \big]}_{n^{\prime}n}^{m^{\prime}m} = \singular_{n + n^{\prime}}^{m - m^{\prime}}, \qquad
    {\big[ \singular\vert\singular \big]}_{n^{\prime}n}^{m^{\prime}m} = \regular_{n^{\prime} - n}^{m - m^{\prime}}
  \end{equation}
\end{subequations}

\noindent
We shall briefly also discuss an alternate, more geometrically inspired mode of translation, namely the method rotation, coäxial translation, and backrotation, or \textsc{rcr} for brevity.
For a truncated multipole expansion, we let not $n$ range to infinity, rather let its greatest extent be $N - 1$.
For some multipole coëfficient, we arrange for each value of $n$ its span in $m$ in an array.
Consider now $\dvec$ given explicitly in terms of spherical coördinates, and the multipole coëfficient be given by $A$.
Say we wish to translate from the singular expansion to an expansion in terms of the regular harmonic, and that this translation matrix $T$ is given such that $A_{\regular} = T_{\singular\vert\regular}(\dvec) A_{\singular}$.
This translation matrix is composed of a rotation matrix $R_{\regular}$, a coäxial displacement matrix $C_{\singular\vert\regular}$, and the inverse rotation matrix $R_{\singular}^{-1}$.
The matter of rotating the spherical harmonics is an involved, but not particularly tedious task.
The rotation will preserve the degree of the spherical harmonic, but it will not preserve the order in general.
That is, should we wish to rotate from the orientation $(\alpha^0, \alpha^1)$ to the orientation $(\beta^0, \beta^1)$, we have that
%
\begin{equation}
  \spherical_n^k(\beta^0, \beta^1) = \sum_{m = -n}^{n} D^n_{mk} \spherical_n^m(\alpha^0, \alpha^1), \qquad
  D^n_{mk} = \exp{(im\alpha^0)} d^n_{mk}(\alpha^1)
\end{equation}

\noindent
Here, $D^n_{mk}$ is the \textsc{Wigner} function, and $d^n_{mk}$ is the corresponding reduced function.%
\footnote{%
  \cite{varsalovic1988quantum} \textsc{Varšalovič}, \textsc{Moskalev}, \& \textsc{Hersonskij}, \S4
}
Thus for each value of $n$ we have one submatrix of size $(2n+1) \times (2n+1)$, accounting for each of the $2n+1$ terms of $m$ and their corresponding map to the same number of rotated terms.
It is clear the rotation matrices will be diagonal, their coëfficient terms corresponding to regularization with respect to the spherical harmonics and either of the regular or singular functions.
Along the diagonal, we then write
%
\begin{subequations}
  \begin{align}
    {\big[ R_{\regular}^n \big]}_{k}^{m} &= \frac{i^{\absl{k}}(n+\absl{m})! c_n^m}{i^{\absl{m}}(n+\absl{k})! c_n^k} D^n_{mk};\\
    %
    {\big[ R_{\singular}^n \big]}_{k}^{m} &= \frac{i^{\absl{m}}(n-\absl{k})! c_n^m}{i^{\absl{k}}(n-\absl{m})! c_n^k} D^n_{mk}.
  \end{align}
\end{subequations}

\noindent
Since the translation $C_{\singular\vert\regular}$ is coäxial, we preserve the azimuthal structure of $m$, and we may thus entirely skip the offset in indexation in \subeqref{\ref{eq:translation_notation}}[b].
The final translation matrix is thus given in its decomposed form; $T_{\singular\vert\regular}(\dvec) = R_{\regular}^{-1}(\alpha^0, \alpha^1) C_{\singular\vert\regular}(d) R_{\singular}(\alpha^0, \alpha^1)$.
In addition to its mathematical elegance, the sparsity of the rotation matrix, and the reduced cost in calculating the coäxial translation should theoretically reduce the needed computational strain from $O(N^4)$ to $O(N^3)$, yielding better performance with proper implementation.
