This work is the culmination of eighteen years of public schooling, five of them in higher education at the University of Oslo.
Thousands of works of just such culminations are carefully crafted each and every year in this very, relatively small city.
Five years is not an insignificant amount of time, and for five years I have chosen to be headstrong and wander off the beaten path at every inpasse.
Although I have meandered, I have not strayed from the goal from the outset of those five years---to learn aerodynamics.

I think my greatest aquired wisdom from all these years of schooling, was conversing with my supervisor, Karsten \textsc{Trulsen}.
I have sometimes entered his office with clear agendas, but often none at all.
Upon us reflecting once on my particular way of studying, he said something along the lines of
\begin{quote}
  ``When I started out as a PhD student, my supervisor CC \textsc{Mei} told me: `Karsten, you need to understand the difference between a \emph{scholar} and a \emph{researcher}.
  When you are a PhD student, you need to be a \emph{researcher}.'
  Simon, you are a \emph{scholar}.''
\end{quote}
My scholarly tendencies have got nothing to do with me not being a PhD student, but rather me always looking back for answers, and in reverence at what has been already done by those that came before.
The researcher looks ahead to new horizons, whilst the humble scholar compiles the findings with his strange Cyrillic characters and forcedly consistent notation.

Although thousands of works just like this one are written each and every year, this one is uniquely mine, and mine alone.
It is the culmination of my meandering with only that one goal in mind---nevermind the path.
I hope my scholarly spirit and ardor for the subjectmatter penetrates its veil of academic dead-pan articulation.
